---
title: "phase2"
author: "Turner Schwiebert"
date: "11/3/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning=F}
library(dplyr)
library(ggplot2)
library(ggthemes)
library(data.table)
library(caret)

forwards <- read.csv("../Data Acquisition/Data/forwards10_20_2022.csv")

attmids_wings <- forwards %>% filter(position == "Att Mid / Wing")
attmids_wings$height_cm <- as.numeric(as.character(attmids_wings$height_cm))

forwards <- forwards %>% filter(position == "Forward")
forwards$height_cm <- as.numeric(as.character(forwards$height_cm))

centerbacks <- read.csv("../Data Acquisition/Data/centerbacks10_20_2022.csv")
centerbacks$height_cm <- as.numeric(as.character(centerbacks$height_cm))

fullbacks <- read.csv("../Data Acquisition/Data/fullbacks10_20_2022.csv")
fullbacks$height_cm <- as.numeric(as.character(fullbacks$height_cm))

midfielders <- read.csv("../Data Acquisition/Data/midfielders10_20_2022.csv")
midfielders$height_cm <- as.numeric(as.character(midfielders$height_cm))
```

# NOTE: the clusters are reproducible, but re-running caret's kmeans changes the order of clusters. Thus, results are reproducible but the order is somewhat random.

# Forwards

## Clustering

```{r, use hierarchical dendrogram to pick number of clusters, only include relevant features}
forward_dendrogram <- hclust(dist(forwards[, c("Aerials.won", "Players.Dribbled.Past", "Touches..Mid.3rd.", "Touches..Att.3rd.", "Crosses", "Shots.Total", "Ball.Recoveries", "Shot.Creating.Actions")]))
plot(forward_dendrogram)
```
3 clusters looks good

```{r, use k-means to create clusters}
forwards_clust <- forwards
kFit = kmeans(forwards[,c("Aerials.won", "Players.Dribbled.Past", "Touches..Mid.3rd.", "Touches..Att.3rd.", "Crosses", "Shots.Total", "Ball.Recoveries", "Shot.Creating.Actions")], 3, nstart=50)
forwards_clust$k_cluster <- kFit$cluster
```

## Validation/Reporting

```{r}
vars = c("Aerials.won", "Players.Dribbled.Past", "Touches..Mid.3rd.", "Touches..Att.3rd.", "Crosses", "Shots.Total", "Ball.Recoveries", "Shot.Creating.Actions")
n_clust = 3
result <- data.frame(cluster = numeric(length(vars)*n_clust),
           variable_name = character(length(vars)*n_clust),
           mean = numeric(length(vars)*n_clust),
           sd = numeric(length(vars)*n_clust),
           total_mean = numeric(length(vars)*n_clust),
           total_sd = numeric(length(vars)*n_clust),
           count = numeric(length(vars)*n_clust))

cur_clust = 1
cur_var = 1

#IF THIS FOR LOOP THROWS AN ERROR, JUST RUN THE CHUNK AGAIN (I dont know why but it works)
for(i in 1:(length(vars)*n_clust)){
  result$cluster[i] = cur_clust
  result$variable_name[i] = vars[cur_var]
  result$mean[i] = round(mean(filter(forwards_clust, k_cluster == cur_clust)[,vars[cur_var]]), 2)
  result$sd[i] = round(sd(filter(forwards_clust, k_cluster == cur_clust)[,vars[cur_var]]), 2)
  result$total_mean[i] = round(mean(forwards_clust[,vars[cur_var]]), 2)
  result$total_sd[i] = round(sd(forwards_clust[,vars[cur_var]]), 2)
  result$count[i] = nrow(filter(forwards_clust, k_cluster == cur_clust))
  cur_var = cur_var + 1
  if(cur_var > length(vars)){
    cur_var = 1
    cur_clust = cur_clust + 1}
}

#view players in individual cluster
filter(forwards_clust, k_cluster == 1)

#insert values into table row by row:
filter(result, variable_name == vars[1])

#view players in individual cluster
result
```

```{r}
ggplot(data = result, aes(x = as.factor(cluster), y = mean, fill = variable_name)) +
  geom_bar(stat = "identity", position = "dodge")+
  theme_minimal()+
  labs(title = "Forwards - Summary Statistics by Cluster", y = "Mean Values", x = "Cluster")
```

# Att Mid/Wings

## Clustering

```{r, use hierarchical dendrogram to pick number of clusters, only include relevant features}
AMW_dendrogram <- hclust(dist(attmids_wings[, c("Players.Dribbled.Past", "Touches..Mid.3rd.", "Touches..Att.3rd.", "Crosses", "Shots.Total", "Ball.Recoveries", "Shot.Creating.Actions", "Progressive.Passes")]))
plot(AMW_dendrogram)
```
4 clusters looks good

```{r, use k-means to create clusters}
AMW_clust <- attmids_wings
kFit = kmeans(attmids_wings[,c("Players.Dribbled.Past", "Touches..Mid.3rd.", "Touches..Att.3rd.", "Crosses", "Shots.Total", "Ball.Recoveries", "Shot.Creating.Actions", "Progressive.Passes")], 4, nstart=50)
AMW_clust$k_cluster <- kFit$cluster
```

## Validation/Reporting

```{r}
vars = c("Players.Dribbled.Past", "Touches..Mid.3rd.", "Touches..Att.3rd.", "Crosses", "Shots.Total", "Ball.Recoveries", "Shot.Creating.Actions", "Progressive.Passes")
n_clust = 4

result <- data.frame(cluster = numeric(length(vars)*n_clust),
           variable_name = character(length(vars)*n_clust),
           mean = numeric(length(vars)*n_clust),
           sd = numeric(length(vars)*n_clust),
           total_mean = numeric(length(vars)*n_clust),
           total_sd = numeric(length(vars)*n_clust),
           count = numeric(length(vars)*n_clust))

cur_clust = 1
cur_var = 1

#IF THIS FOR LOOP THROWS AN ERROR, JUST RUN THE CHUNK AGAIN (I dont know why but it works)
for(i in 1:(length(vars)*n_clust)){
  result$cluster[i] = cur_clust
  result$variable_name[i] = vars[cur_var]
  result$mean[i] = round(mean(filter(AMW_clust, k_cluster == cur_clust)[,vars[cur_var]]), 2)
  result$sd[i] = round(sd(filter(AMW_clust, k_cluster == cur_clust)[,vars[cur_var]]), 2)
  result$total_mean[i] = round(mean(AMW_clust[,vars[cur_var]]), 2)
  result$total_sd[i] = round(sd(AMW_clust[,vars[cur_var]]), 2)
  result$count[i] = nrow(filter(AMW_clust, k_cluster == cur_clust))
  cur_var = cur_var + 1
  if(cur_var > length(vars)){
    cur_var = 1
    cur_clust = cur_clust + 1}
}

#view players in individual cluster
filter(AMW_clust, k_cluster == 3)

#insert values into table:
filter(result, variable_name == vars[1])

#analyze resulting summary statistics
result
```

```{r}
ggplot(data = result, aes(x = as.factor(cluster), y = mean, fill = variable_name)) +
  geom_bar(stat = "identity", position = "dodge")+
  theme_minimal()+
  labs(title = "Attacking Mids/Wingers - Summary Statistics by Cluster", y = "Mean Values", x = "Cluster")
```

# Midfielders

## Clustering

```{r, use hierarchical dendrogram to pick number of clusters, only include relevant features}
Mid_dendrogram <- hclust(dist(midfielders[, c("Players.Dribbled.Past", "Touches..Def.3rd.", "Touches..Mid.3rd.", "Touches..Att.3rd.", "Crosses", "Shots.Total", "Ball.Recoveries", "Shot.Creating.Actions", "Progressive.Passes")]))
plot(Mid_dendrogram)
```
5 clusters looks good, midfielders are used many ways in the sport

```{r, use k-means to create clusters}
Mid_clust <- midfielders
kFit = kmeans(Mid_clust[,c("Players.Dribbled.Past", "Touches..Def.3rd.", "Touches..Mid.3rd.", "Touches..Att.3rd.", "Crosses", "Shots.Total", "Ball.Recoveries", "Shot.Creating.Actions", "Progressive.Passes")], 5, nstart=50)
Mid_clust$k_cluster <- kFit$cluster
```

## Validation/Reporting

```{r}
vars = c("Players.Dribbled.Past", "Touches..Def.3rd.", "Touches..Mid.3rd.", "Touches..Att.3rd.", "Crosses", "Shots.Total", "Ball.Recoveries", "Shot.Creating.Actions", "Progressive.Passes")
n_clust = 5

result <- data.frame(cluster = numeric(length(vars)*n_clust),
           variable_name = character(length(vars)*n_clust),
           mean = numeric(length(vars)*n_clust),
           sd = numeric(length(vars)*n_clust),
           total_mean = numeric(length(vars)*n_clust),
           total_sd = numeric(length(vars)*n_clust),
           count = numeric(length(vars)*n_clust))

cur_clust = 1
cur_var = 1

#IF THIS FOR LOOP THROWS AN ERROR, JUST RUN THE CHUNK AGAIN (I dont know why but it works)
for(i in 1:(length(vars)*n_clust)){
  result$cluster[i] = cur_clust
  result$variable_name[i] = vars[cur_var]
  result$mean[i] = round(mean(filter(Mid_clust, k_cluster == cur_clust)[,vars[cur_var]]), 2)
  result$sd[i] = round(sd(filter(Mid_clust, k_cluster == cur_clust)[,vars[cur_var]]), 2)
  result$total_mean[i] = round(mean(Mid_clust[,vars[cur_var]]), 2)
  result$total_sd[i] = round(sd(Mid_clust[,vars[cur_var]]), 2)
  result$count[i] = nrow(filter(Mid_clust, k_cluster == cur_clust))
  cur_var = cur_var + 1
  if(cur_var > length(vars)){
    cur_var = 1
    cur_clust = cur_clust + 1}
}

#view players in individual cluster
filter(Mid_clust, k_cluster == 1)

#insert values into table row by row:
filter(result, variable_name == vars[9])

#analyze table of resulting summary statistics
result
```

```{r}
ggplot(data = result, aes(x = as.factor(cluster), y = mean, fill = variable_name)) +
  geom_bar(stat = "identity", position = "dodge")+
  theme_minimal()+
  labs(title = "Midfielders - Summary Statistics by Cluster", y = "Mean Values", x = "Cluster")
```

# Fullbacks

## Clustering

```{r, use hierarchical dendrogram to pick number of clusters, only include relevant features}
FullB_dendrogram <- hclust(dist(fullbacks[, c("Players.Dribbled.Past", "Touches..Def.3rd.", "Touches..Mid.3rd.", "Touches..Att.3rd.", "Crosses", "Shots.Total", "Ball.Recoveries", "Shot.Creating.Actions", "Progressive.Passes")]))
plot(FullB_dendrogram)
```
4 clusters looks good

```{r, use k-means to create clusters}
fullB_clust <- fullbacks
kFit = kmeans(fullB_clust[,c("Players.Dribbled.Past", "Touches..Def.3rd.", "Touches..Mid.3rd.", "Touches..Att.3rd.", "Crosses", "Shots.Total", "Ball.Recoveries", "Shot.Creating.Actions", "Progressive.Passes")], 4, nstart=50)
fullB_clust$k_cluster <- kFit$cluster
```

## Validation/Reporting

```{r}
vars = c("Players.Dribbled.Past", "Touches..Def.3rd.", "Touches..Mid.3rd.", "Touches..Att.3rd.", "Crosses", "Shots.Total", "Ball.Recoveries", "Shot.Creating.Actions", "Progressive.Passes")
n_clust = 4
result <- data.frame(cluster = numeric(length(vars)*n_clust),
           variable_name = character(length(vars)*n_clust),
           mean = numeric(length(vars)*n_clust),
           sd = numeric(length(vars)*n_clust),
           total_mean = numeric(length(vars)*n_clust),
           total_sd = numeric(length(vars)*n_clust),
           count = numeric(length(vars)*n_clust))

cur_clust = 1
cur_var = 1

#IF THIS FOR LOOP THROWS AN ERROR, JUST RUN THE CHUNK AGAIN (I dont know why but it works)
for(i in 1:(length(vars)*n_clust)){
  result$cluster[i] = cur_clust
  result$variable_name[i] = vars[cur_var]
  result$mean[i] = round(mean(filter(fullB_clust, k_cluster == cur_clust)[,vars[cur_var]]), 2)
  result$sd[i] = round(sd(filter(fullB_clust, k_cluster == cur_clust)[,vars[cur_var]]), 2)
  result$total_mean[i] = round(mean(fullB_clust[,vars[cur_var]]), 2)
  result$total_sd[i] = round(sd(fullB_clust[,vars[cur_var]]), 2)
  result$count[i] = nrow(filter(fullB_clust, k_cluster == cur_clust))
  cur_var = cur_var + 1
  if(cur_var > length(vars)){
    cur_var = 1
    cur_clust = cur_clust + 1}
}

#view players in individual cluster
filter(fullB_clust, k_cluster == 1)

#insert values into table row by row:
filter(result, variable_name == vars[9])

#analyze table of resulting summary statistics
result
```

```{r}
ggplot(data = result, aes(x = as.factor(cluster), y = mean, fill = variable_name)) +
  geom_bar(stat = "identity", position = "dodge")+
  theme_minimal()+
  labs(title = "Fullbacks - Summary Statistics by Cluster", y = "Mean Values", x = "Cluster")
```

# Centerbacks

## Clustering

```{r, use hierarchical dendrogram to pick number of clusters, only include relevant features}
CB_dendrogram <- hclust(dist(centerbacks[, c("Aerials.won", "Touches..Def.3rd.", "Touches..Mid.3rd.", "Touches..Att.3rd.", "Shots.Total", "Ball.Recoveries", "Progressive.Passes")]))
plot(CB_dendrogram)
```
2 clusters looks good, centerbacks are only used defensively in today's game

```{r, use k-means to create clusters}
CB_clust <- centerbacks
kFit = kmeans(CB_clust[,c("Aerials.won", "Touches..Def.3rd.", "Touches..Mid.3rd.", "Touches..Att.3rd.", "Shots.Total", "Ball.Recoveries", "Progressive.Passes")], 2, nstart=50)
CB_clust$k_cluster <- kFit$cluster
```

## Validation/Reporting

```{r}
vars = c("Aerials.won", "Touches..Def.3rd.", "Touches..Mid.3rd.", "Touches..Att.3rd.", "Shots.Total", "Ball.Recoveries", "Progressive.Passes")
n_clust = 2
result <- data.frame(cluster = numeric(length(vars)*n_clust),
           variable_name = character(length(vars)*n_clust),
           mean = numeric(length(vars)*n_clust),
           sd = numeric(length(vars)*n_clust),
           total_mean = numeric(length(vars)*n_clust),
           total_sd = numeric(length(vars)*n_clust),
           count = numeric(length(vars)*n_clust))

cur_clust = 1
cur_var = 1

#IF THIS FOR LOOP THROWS AN ERROR, JUST RUN THE CHUNK AGAIN (I dont know why but it works)
for(i in 1:(length(vars)*n_clust)){
  result$cluster[i] = cur_clust
  result$variable_name[i] = vars[cur_var]
  result$mean[i] = round(mean(filter(CB_clust, k_cluster == cur_clust)[,vars[cur_var]]), 2)
  result$sd[i] = round(sd(filter(CB_clust, k_cluster == cur_clust)[,vars[cur_var]]), 2)
  result$total_mean[i] = round(mean(CB_clust[,vars[cur_var]]), 2)
  result$total_sd[i] = round(sd(CB_clust[,vars[cur_var]]), 2)
  result$count[i] = nrow(filter(CB_clust, k_cluster == cur_clust))
  cur_var = cur_var + 1
  if(cur_var > length(vars)){
    cur_var = 1
    cur_clust = cur_clust + 1}
}

#view players in individual cluster
filter(CB_clust, k_cluster == 1)

#insert values into table row by row:
filter(result, variable_name == vars[1])

#analyze table of resulting summary statistics
result
```

```{r}
ggplot(data = result, aes(x = as.factor(cluster), y = mean, fill = variable_name)) +
  geom_bar(stat = "identity", position = "dodge")+
  theme_minimal()+
  labs(title = "Center Backs - Summary Statistics by Cluster", y = "Mean Values", x = "Cluster")
```











